---
title: "Mycorrhizal fungi affect the proportions and chemistry of different soil organic matter fractions"
---

<style type="text/css">
.table {

    width: 50%;
    align: 'center';

}
</style>

```{r setup, include=FALSE}

library(tidyverse)
library(knitr)
library(ggpubr)
library(cowplot)
library(ggmap)
library(maps)
library(mapdata)
library(stringi)
library(nlme)
library(ggsn)
library(sf)
library(car)
library(plotly)

#bringing in foliar chemistry data from Northern Ecosystems Research Cooperative: http://www.nercscience.org/Metadata_FoliarChemistry.html
nerc=read.csv("~/Documents/Projects/inProgress/Myc_MAOM/data/foliar_chem/FoliarChem_NERC.csv", header=TRUE) %>% 
  dplyr::select(SpeciesID, StateOrProvince, Foliage_cellulose_., Foliage_lignin_., Foliage_C_., Foliage_N_.)  %>%
  group_by(SpeciesID) %>% 
  mutate(C.N= Foliage_C_. / Foliage_N_.,
         lignin.N=Foliage_lignin_. / Foliage_N_.,
         se.c.n=(sd(C.N, na.rm=TRUE) /  
           sqrt(length(C.N[!is.na(C.N)]))),
         n.C.N.values=length(C.N[!is.na(C.N)]),
         se.lig.n=(sd(lignin.N, na.rm=TRUE) /  
           sqrt(length(lignin.N[!is.na(lignin.N)]))),
         n.Lig.N.values=length(lignin.N[!is.na(lignin.N)])) %>%
  filter(SpeciesID=="THOC2" |
           SpeciesID=="TSCA" |
           SpeciesID=="FAGR" |
           SpeciesID=="ACSA3" |
           SpeciesID=="FRAM2" |
           SpeciesID=="BEAL2") %>% 
  summarise(avg_CN=mean(C.N, na.rm=TRUE),
            avg_lig.N=mean(lignin.N, na.rm=TRUE),
            se_CN=mean(se.c.n, na.rm=TRUE),
            se_lig.N=mean(se.lig.n, na.rm=TRUE),
            C.N.count=mean(n.C.N.values),
            Lig.N.count=mean(n.Lig.N.values)) %>% 
  mutate(SpeciesID=case_when(SpeciesID=="THOC2" ~ "T. occidentalis",
                             SpeciesID=="TSCA" ~ "T. canadensis",
                             SpeciesID=="FAGR" ~ "F. grandifolia",
                             SpeciesID=="ACSA3" ~ "A. saccharum",
                             SpeciesID=="FRAM2" ~ "F. americana",
                             SpeciesID=="BEAL2" ~ "B. alleghaniensis"),
         Myc=case_when(SpeciesID=="T. occidentalis" ~ "AM",
                       SpeciesID=="A. saccharum" ~ "AM",
                       SpeciesID=="F. americana" ~ "AM",
                       SpeciesID=="T. canadensis" ~ "ECM",
                       SpeciesID=="F. grandifolia" ~ "ECM",
                       SpeciesID=="B. alleghaniensis" ~ "ECM"))

nerc$species= factor(nerc$SpeciesID, levels=c("F. americana", "A. saccharum", "T. occidentalis", "B. alleghaniensis", "F. grandifolia", "T. canadensis"))



#This is site-level information about locations where samples were collected and soil stuff
sites=read.csv("~/Documents/Projects/inProgress/Myc_MAOM/Data/Plot Data/MycMin_PlotData.csv") %>% 
  dplyr::select(Site, Target.species, Latitude, Longitude, Soil.clay, Soil.silt, Soil.sand) %>% 
  mutate(Latitude=as.character(Latitude),
         Longitude=as.character(Longitude),
         site=case_when(Site=="RR" ~ "CV",
                        Site=="CB" ~ "CV",
                        Site=="AB" ~ "AB",
                        Site=="HB" ~ "HB",
                        Site=="MK" ~ "MK",
                        Site=="MS" ~ "MS",
                        Site=="RM" ~ "RM",
                        Site=="LP" ~ "LP")) %>% 
  group_by(site) %>% 
  summarise(Lat=mean(as.numeric(Latitude)),
            Long=mean(as.numeric(Longitude)),
            clay=mean(Soil.clay),
            silt=mean(Soil.silt),
            sand=mean(Soil.sand))


#Here's my data about soil C and N in the different fractions and in bulk samples
myc.maom=read.csv("~/Documents/Projects/inProgress/Myc_MAOM/data/myc.maom.data.csv") %>%
  mutate(species=case_when(Species=="ACSA" ~ "A. saccharum",
                           Species=="FRAM" ~ "F. americana",
                           Species=="FAGR" ~ "F. grandifolia",
                           Species=="BEAL" ~ "B. alleghaniensis",
                           Species=="THOC" ~ "T. occidentalis",
                           Species=="TSCA" ~ "T. canadensis"),
         gC.as.FLF=((FLF.pct.C/100)*FLF.g)*(Sample.mass/20),
         gC.as.OLF=((OLF.pct.C/100)*OLF.g)*(Sample.mass/20),
         gC.as.HF=((HF.pct.C/100)*HF.g)*(Sample.mass/20),
         gN.as.FLF=((FLF.pct.N/100)*FLF.g)*(Sample.mass/20),
         gN.as.OLF=((OLF.pct.N/100)*OLF.g)*(Sample.mass/20),
         gN.as.HF=((HF.pct.N/100)*HF.g)*(Sample.mass/20),
         org.CN=org.pct.C/org.pct.N,
         min.CN=min.pct.C/min.pct.N)%>% 
  mutate(propC.as.FLF=gC.as.FLF/(gC.as.FLF+gC.as.OLF+gC.as.HF),
         propC.as.OLF=gC.as.OLF/(gC.as.FLF+gC.as.OLF+gC.as.HF),
         propC.as.HF=gC.as.HF/(gC.as.FLF+gC.as.OLF+gC.as.HF),
         propN.as.FLF=gN.as.FLF/(gN.as.FLF+gN.as.OLF+gN.as.HF),
         propN.as.OLF=gN.as.OLF/(gN.as.FLF+gN.as.OLF+gN.as.HF),
         propN.as.HF=gN.as.HF/(gN.as.FLF+gN.as.OLF+gN.as.HF),
         FLF.mgC.per.g.soil=gC.as.FLF*(1000)/Sample.mass,
         OLF.mgC.per.g.soil=gC.as.OLF*(1000)/Sample.mass,
         HF.mgC.per.g.soil=gC.as.HF*(1000)/Sample.mass,
         FLF.mgN.per.g.soil=gN.as.FLF*(1000)/Sample.mass,
         OLF.mgN.per.g.soil=gN.as.OLF*(1000)/Sample.mass,
         HF.mgN.per.g.soil=gN.as.HF*(1000)/Sample.mass,
         site=case_when(Site=="RR" ~ "CV",
                        Site=="CB" ~ "CV",
                        Site=="AB" ~ "AB",
                        Site=="HB" ~ "HB",
                        Site=="MK" ~ "MK",
                        Site=="MS" ~ "MS",
                        Site=="RM" ~ "RM",
                        Site=="LP" ~ "LP")) %>% 
  left_join(nerc, by= "species") %>% 
  dplyr::select(-SpeciesID, -Myc.y, -Species) %>% 
    left_join(sites, by="site")


#ordering species how i want them to appear in figures
myc.maom$species= factor(myc.maom$species, levels=c("F. americana", "A. saccharum", "T. occidentalis", "B. alleghaniensis", "F. grandifolia", "T. canadensis"))


# #Long version of this data frame...
myc.maom.long=read.csv("~/Documents/Projects/inProgress/Myc_MAOM/data/myc.maom.data.csv") %>%
  dplyr::select(-X.1,-X) %>%
  mutate(gC.as.FLF=((FLF.pct.C/100)*FLF.g),
         gC.as.OLF=((OLF.pct.C/100)*OLF.g),
         gC.as.HF=((HF.pct.C/100)*HF.g),
         gN.as.FLF=((FLF.pct.N/100)*FLF.g),
         gN.as.OLF=((OLF.pct.N/100)*OLF.g),
         gN.as.HF=((HF.pct.N/100)*HF.g),
         org.CN=org.pct.C/org.pct.N,
         min.CN=min.pct.C/min.pct.N) %>%
  mutate(propC.as.FLF=gC.as.FLF/(gC.as.FLF+gC.as.OLF+gC.as.HF),
         propC.as.OLF=gC.as.OLF/(gC.as.FLF+gC.as.OLF+gC.as.HF),
         propC.as.HF=gC.as.HF/(gC.as.FLF+gC.as.OLF+gC.as.HF),
         propN.as.FLF=gN.as.FLF/(gN.as.FLF+gN.as.OLF+gN.as.HF),
         propN.as.OLF=gN.as.OLF/(gN.as.FLF+gN.as.OLF+gN.as.HF),
         propN.as.HF=gN.as.HF/(gN.as.FLF+gN.as.OLF+gN.as.HF),
         FLF.mgC.per.g.soil=gC.as.FLF*(1000)/Sample.mass,
         OLF.mgC.per.g.soil=gC.as.OLF*(1000)/Sample.mass,
         HF.mgC.per.g.soil=gC.as.HF*(1000)/Sample.mass,
         FLF.mgN.per.g.soil=gN.as.FLF*(1000)/Sample.mass,
         OLF.mgN.per.g.soil=gN.as.OLF*(1000)/Sample.mass,
         HF.mgN.per.g.soil=gN.as.HF*(1000)/Sample.mass) %>%
  pivot_longer(-c(uid,Site,Species,Rep,Myc,Sample.mass,O.horizon.depth), names_to= "variable", values_to="value.as.fraction") %>%
  mutate(fraction = case_when(stri_detect_fixed(variable, "FLF") ~ "FLF",
                              stri_detect_fixed(variable, "OLF")    ~ "OLF",
                              stri_detect_fixed(variable, "HF")    ~ "HF",
                              TRUE  ~ "Bulk"),
         type= case_when(stri_detect_fixed(variable, "pct") ~ "percent",
                         stri_detect_fixed(variable, ".mg") ~ "milligrams",
                         stri_detect_fixed(variable, "prop") ~ "proportion",
                         TRUE ~ "other"),
         element= case_when(stri_detect_fixed(variable, "C") ~ "carbon",
                            stri_detect_fixed(variable, "N") ~ "nitrogen",
                            TRUE ~ "soil"),
         measurement = paste(type, element, sep="_"),
         species= case_when(Species=="ACSA" ~ "A. saccharum",
                           Species=="FRAM" ~ "F. americana",
                           Species=="FAGR" ~ "F. grandifolia",
                           Species=="BEAL" ~ "B. alleghaniensis",
                           Species=="THOC" ~ "T. occidentalis",
                           Species=="TSCA" ~ "T. canadensis"),
         site=case_when(Site=="RR" ~ "CV",
                        Site=="CB" ~ "CV",
                        Site=="AB" ~ "AB",
                        Site=="HB" ~ "HB",
                        Site=="MK" ~ "MK",
                        Site=="MS" ~ "MS",
                        Site=="RM" ~ "RM",
                        Site=="LP" ~ "LP")) %>%
  left_join(nerc, by= "species") %>%
  dplyr::select(-SpeciesID, -Myc.y, -Species) %>%
    left_join(sites, by="site")


# myc.maom.mgC=myc.maom.long %>% 
#   filter(measurement=="milligrams_carbon",
#          fraction != "OLF") 
# 
# myc.maom.propC=myc.maom.long %>% 
#   filter(measurement=="proportion_carbon",
#          fraction != "OLF")
# 
# myc.maom.mgN=myc.maom.long %>% 
#   filter(measurement=="milligrams_nitrogen",
#          fraction != "OLF") 
# 
# myc.maom.propN=myc.maom.long %>% 
#   filter(measurement=="proportion_nitrogen",
#          fraction != "OLF")

site.min.N=myc.maom %>% 
  group_by(Site) %>% 
  summarise(site.N=mean(min.pct.N))

#These are color palettes:
ExCon= c('#CCBB44', '#114C3A')
cbbPalette <- c("#000000", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
pretty_pal=c("#973204", "#C2956E", "#264737", "#4ca489","gray32",  "gray81", "#441507")
pretty_pal2=c("#4ca489", "#264737", "darkseagreen1")


```

****

## Quick Summary
### *Recent studies indicate that ecosystems dominated by vegetation associated with AM fungi store more soil organic matter (SOM) as mineral-associated organic matter (MAOM) rather than particulate organic matter (POM), but plant mycorrhizal associations are often correlated with plant litter chemistry. Because plant litter chemistry is also known to influence MAOM formation, is is difficult to disentangle these potential drivers of MAOM content. We  in forests across New Hampshire and Vermont.*  



**But what are MAOM and POM?**

Mineral-associated organic matter (MAOM) is increasingly recognized as a stable and long-term pool of soil carbon and nutrients. Recent work suggests MAOM is primarily of microbial origin, composed of microbial necromass and highly decomposed organic matter residues. These low-molecular weight organic compounds form strong bonds with mineral surfaces, particularly clays and metal oxides. Because these organo-mineral bonds restrict the access of decomposers to MAOM molecules, this material may persist in soil for centuries to millennia. Further, because its long-term stability is based on physical barriers that prevent microbial access rather than energetic constraints on decomposers, MAOM may be less susceptible to decay with increasing soil temperature.

Particulate organic matter (POM) is more or less the "bits" of leaves and roots that remain relatively undecomposed in soil. You can think of POM as organic matter that is not quite done breaking down. Sometimes this decomposition takes a long time, and carbon and nutrients can be stored as POM for years. This is especially true when the chemistry of the original material is complex (in other words, hard for microbes to break apart). Wood, pine needles, thick, waxy leaves, and leaves with a lot of lignin may be considered complex litter material, while thinner, more nitrogen-rich leaves are less complex and easier for microbes to break down.

Based on this information, we may expect litter quality to affect the amount of POM in soil. But things get a little more complicated when we think about the formation of MAOM. Since MAOM is organic matter that has traveled further along in the decomposition process, it is the product of all the microbial activity that modified it in these various stages of decay. In short, the more *efficiently* the microbial community can decompose organic matter, the more microbial biomass is generated in the soil, and the faster the production of the organic residues and microbial necromass that are the putative precursors of MAOM. Certain microbes are more efficient at decomposition than others, and microbes are all generally more efficient decomposers when the chemistry of the organic matter they are using is less complex. Therefore, MAOM formation depends both on the activity and composition of the microbes conducting decomposition as well as the chemistry of the organic matter available for decomposition.

**To recap: MAOM formation is probably affected by...**   
1. The quality of the litter inputs to the soil  
2. The microbial community in the soil

*Other controls that are important to consider:  
1. Mineral composition of the soil in each site (aka what minerals are available for MAOM to stick to!)  
2. Other traits of the trees themselves (root turnover rate, above:belowground biomass, etc)*


#### In this study, we investigated how mycorrhizal associations and foliar chemsitry affect soil MAOM content by measuring MAOM and POM content beneath six tree species carefully selected to represent a range of both factors. We are using the tree's mycorrhizal association as a proxy for the general composition of the microbial communities in the soil.  

### The cast of characters: Our 6 Tree Species  

| Litter Quality |        AM            |        ECM      |
| -------------  |:--------------------:| :--------------:|
| High           | White Ash            |    Yellow Birch |
| Medium         | Sugar Maple          | American Beech  |
| Low            | Northern White Cedar | Eastern Hemlock |  


## We can see the variation in foliar chemistry within each mycorrhizal classification:  

```{r ligninN, echo=FALSE, warning=F, message=F, fig.width=7}

littercn=ggplot(data=nerc, aes(x=species, y=avg_CN, fill=Myc))+
   geom_col()+
  scale_fill_manual(name="Mycorrhizal\nType",values= pretty_pal2)+
  labs(x=" ", y="Foliar C:N")+
  theme_cowplot()+
  theme(legend.position = "right", legend.title=element_text(hjust=0.5),axis.text.x = element_text(face="italic", angle = 30, hjust=1, vjust=1))+
  geom_errorbar(aes(ymin=avg_CN-se_CN, ymax=avg_CN+se_CN), width=0.2, size=0.25, colour="black", linetype="solid")
littercn
ggsave("litterCN.jpg", path="~/Documents/Projects/inProgress/Myc_MAOM/figures", width= 16, height= 14, units="cm")

litterlign=ggplot(data=nerc, aes(x=species, y=avg_lig.N, fill=Myc))+
  geom_col()+
    scale_fill_manual(values= pretty_pal2)+
  labs(x=" ", y="Foliar Lignin:N")+
  theme_cowplot()+
  geom_errorbar(aes(ymin=avg_lig.N-se_lig.N, ymax=avg_lig.N+se_lig.N), width=0.2, size=0.25, colour="black", linetype="solid")


litter=ggarrange(littercn, litterlign, ncol=2)

```

## And to orient yourself, here is where this all took place. We measured MAOM and POM beneath these trees at a variety of sites to make sure we captured the full range of soil contditions that may be found where these species tend to grow.  

```{r map, echo=FALSE, warning=FALSE, message=FALSE}

states <- map_data("state")
nhvt <- subset(states, region %in% c("new hampshire", "vermont"))

regionalmap=ggplot(data=nhvt)+
  geom_polygon(aes(x=long, y=lat, group=group), fill="darkseagreen3", color="black")+
  coord_fixed(1.3)+
  theme(axis.text=element_blank(), axis.line = element_blank(), axis.ticks=element_blank(), axis.title=element_blank(), panel.background = element_rect(fill = "lightgray"), panel.border = element_rect(linetype="solid", fill=NA))+
  geom_point(aes(x = Long, y = Lat), data = sites, size = 1, color="black")+
  geom_rect(aes(xmin = -72.6, xmax = -71.5, ymin = 43.5, ymax = 44.75),
               fill = "transparent", color = "black")

closeUpmap=ggplot(data=nhvt)+
  geom_polygon(aes(x=long, y=lat, group=group), fill="darkseagreen1", color="black")+
  labs(x="Longitude", y="Latitude")+
  coord_fixed(xlim = c(-72.6, -71.5),  ylim = c(43.5, 44.75), ratio=1.3)+
  geom_point(aes(x = Long, y = Lat, label=site), data = sites, size = 3, color="black")+
  geom_text(data = sites, aes(x = Long, y = Lat, label = site), size = 5, vjust = 1.5, hjust = 0.5, colour="black")+
  scalebar(x.min=-72.6, x.max=-71.6, y.min=43.5, y.max=44.75, dist = 20, dist_unit = "km", model = "WGS84", transform=TRUE)

  
gg_inset_map1 = ggdraw() +
  draw_plot(closeUpmap) +
  draw_plot(regionalmap, x = 0.55, y = 0.7, width = 0.3, height = 0.3)

  
gg_inset_map1
```


### Soil texture varied from site to site, with some sites having more clay than others:  
```{r soiltexture, echo=FALSE, warning=FALSE, message=FALSE, fig.width=12}

clay=ggplot(sites, aes(x=site, y=clay))+geom_bar(stat="identity", fill="#B9684A")+
  labs(x="Site", y="% Clay")
silt=ggplot(sites, aes(x=site, y=silt))+geom_bar(stat="identity", fill="#C2956E")+
  labs(x="Site", y="% Silt")
sand=ggplot(sites, aes(x=site, y=sand))+geom_bar(stat="identity", fill="#E5D1B8")+
  labs(x="Site", y="% Sand")

tex1=ggarrange(sand,silt,clay,ncol=3, nrow=1)
tex1

```  


****
## Statistical Analysis:  
### We tested how mycorrhizal associations, foliar chemistry, and clay content affected the proportion of total soil carbon found in the MAOM vs POM, and the C:N of both the MAOM and POM with a series of relatively simple regressions:  
|Model #  | Response Variables   |  Fixed effects                               |  
|:-------:| ---------------------|:--------------------------------------------:|
|1        | proportion C in MAOM | Mycorrhizal type<br>Foliar C:N<br>Soil%clay  | 
|2        | proportion C in POM  | Mycorrhizal type<br>Foliar C:N<br>Soil%clay  |
|3        | MAOM C:N             | Mycorrhizal type<br>Foliar C:N<br>Soil%clay  | 
|4        | POM C:N              | Mycorrhizal type<br>Foliar C:N<br>Soil%clay  | 
Remember: % Clay was determined at the site level; there are 7 distinct sites. Foliar C:N was determined at the species level; there are 6 distinct species.

***  

## Results: 
## Let's start with the relative amount of carbon stored in each organic matter fraction.  

### 1. The proportion of total C in the MAOM fraction was lower in ECM plots and increased with soil clay percentage.   
 
### 2. The proportion of total C in the POM fraction was higher under trees with high foliar C:N and decreased with higher soil clay percentage.  

```{r figure_propC, echo=FALSE, warning=FALSE, message=FALSE, fig.width=20, fig.height=14}

HF.propC.by.clay=ggplot(myc.maom, aes(x=clay, y=propC.as.HF, colour=Myc.x))+
   geom_smooth(method="lm",size=2)+
   geom_point(size=3, alpha=0.5)+
  scale_colour_manual(values=pretty_pal2)+
  labs(x=" % Clay ", y="MAOM C (proportion)", colour="Mycorrhizal\nType")+
  theme_cowplot()+
  theme(legend.title=element_text(hjust=0.5))

HF.propC.by.cn=ggplot(myc.maom, aes(x=avg_CN, y=propC.as.HF, colour=Myc.x))+
   geom_smooth(method="lm",size=2, linetype="dashed")+
   geom_point(size=3, alpha=0.5)+
  scale_colour_manual(values=pretty_pal2)+
  labs(x=" Foliar C:N ", y="MAOM C (proportion)", colour="Mycorrhizal\nType")+
  theme_cowplot()+
  theme(legend.title=element_text(hjust=0.5))

FLF.propC.by.clay=ggplot(myc.maom, aes(x=clay, y=propC.as.FLF, colour=Myc.x))+
   geom_smooth(method="lm",size=2)+
   geom_point(size=3, alpha=0.5)+
  scale_colour_manual(values=pretty_pal2)+
  labs(x=" % Clay ", y="POM C (proportion)", colour="Mycorrhizal\nType")+
  theme_cowplot()+
  theme(legend.title=element_text(hjust=0.5))

FLF.propC.by.cn=ggplot(myc.maom, aes(x=avg_CN, y=propC.as.FLF, colour=Myc.x))+
   geom_smooth(method="lm",size=2)+
   geom_point(size=3, alpha=0.5)+
  scale_colour_manual(values=pretty_pal2)+
  labs(x=" Foliar C:N ", y="POM C (proportion)", colour="Mycorrhizal\nType")+
  theme_cowplot()+
  theme(legend.title=element_text(hjust=0.5))


figure.propC=ggarrange(HF.propC.by.cn, HF.propC.by.clay, FLF.propC.by.cn,  FLF.propC.by.clay, nrow=2, ncol=2, common.legend = TRUE,legend = "right")
figure.propC
ggsave("propC.jpg", path="~/Documents/Projects/inProgress/Myc_MAOM/figures", scale=1, width= 21, height= 18, units="cm")


```

***  
## Now to look at the chemistry of these organic matter fractions:
### 1. MAOM C:N was lower in AM-dominated stands, lower where there was more clay, and higher where the litter was nasty.  
```{r CN_MAOM,  echo=FALSE, warning=FALSE, message=FALSE, fig.width=10, include=FALSE}
#Need to massage this df a bit...
myc.maom.cn=myc.maom %>% 
  dplyr::select(uid, Myc.x, clay, avg_CN, FLF.pct.C, FLF.pct.N, HF.pct.C, HF.pct.N) %>% 
  mutate(FLF.cn= FLF.pct.C/FLF.pct.N,
         HF.cn=HF.pct.C/HF.pct.N) 

som.cn.summary=myc.maom.cn %>% 
  group_by(Myc.x) %>% 
  summarise(mean.maom.cn=mean(HF.cn),
            mean.POM.cn=mean(FLF.cn[!is.na(FLF.cn)]))

myc.maom.cn.long=myc.maom.cn%>% 
   pivot_longer(c(FLF.cn, HF.cn), names_to= "Fraction", values_to="C.N") %>%
   mutate(fraction = case_when(stri_detect_fixed(Fraction, "FLF") ~ "FLF",
                               TRUE  ~ "HF"))


cn_mod_hf=lm(HF.cn ~ Myc.x + avg_CN + clay, data=myc.maom.cn, na.action=na.omit)

summary(cn_mod_hf)
```

### 2. POM C:N followed foliar C:N (this is likely because it still *is* basically litter... and no effect of mycorrhizal type on POM C:N is evidence that our study design was effective!)   

```{r fig4, echo=FALSE, warning=FALSE, message=FALSE, fig.width=20, fig.height=16}
maom.cn.vs.litter=ggplot(myc.maom.cn, aes(x=avg_CN, y=HF.cn, colour=Myc.x))+
    geom_smooth(method="lm",size=2)+
   geom_point(size=3, alpha=0.5)+
   scale_colour_manual(values=pretty_pal2)+
   labs(x="Foliar C:N ", y="MAOM C:N", colour="Mycorrhizal\nType")+
   theme_cowplot()+
  theme(legend.title=element_text(hjust=0.5))
 
maom.cn.vs.clay=ggplot(myc.maom.cn, aes(x=clay, y=HF.cn, colour=Myc.x))+
    geom_smooth(method="lm",size=2)+
   geom_point(size=3, alpha=0.5)+
   scale_colour_manual(values=pretty_pal2)+
   labs(x=" % Clay ", y="MAOM C:N", colour="Mycorrhizal\nType")+
   theme_cowplot()+
  theme(legend.title=element_text(hjust=0.5))

pom.cn.vs.litter=ggplot(myc.maom.cn, aes(x=avg_CN, y=FLF.cn, colour=Myc.x))+
    geom_smooth(method="lm",size=2)+
   geom_point(size=3, alpha=0.5)+
   scale_colour_manual(values=pretty_pal2)+
   labs(x="Foliar C:N ", y="POM C:N", colour="Mycorrhizal\nType")+
   theme_cowplot()+
  theme(legend.title=element_text(hjust=0.5))

pom.cn.vs.clay=ggplot(myc.maom.cn, aes(x=clay, y=FLF.cn, colour=Myc.x))+
    geom_smooth(method="lm",size=2)+
   geom_point(size=3, alpha=0.5)+
   scale_colour_manual(values=pretty_pal2)+
   labs(x=" % Clay ", y="POM C:N", colour="Mycorrhizal\nType")+
   theme_cowplot()+
  theme(legend.title=element_text(hjust=0.5))


figure.maom.pom.cn=ggarrange(maom.cn.vs.litter, maom.cn.vs.clay, pom.cn.vs.litter, pom.cn.vs.clay, nrow=2, ncol=2, common.legend=TRUE,legend = "right")
figure.maom.pom.cn
ggsave("cn.jpg", path="~/Documents/Projects/inProgress/Myc_MAOM/figures", width= 21, height= 18, units="cm")



```

# Take-home message: More of the C in soil beneath AM-associated trees (vs. ECM-associated trees) becomes associated with minerals in soil, even when the chemistry of the incoming litter is really complex (like cedar leaves!). This means AM-dominated forests may send less soil C to the atmosphere than ECM-dominated forests as the earth gets warmer (and in the case of New England, wetter). This stable pool of MAOM is also higher in nitrogen than the MAOM beneath ECM trees, which may come in handy as a reserve of soil N to fuel plant and microbial productivity. These data support the use of mycorrhizal associations as predictors of soil organic matter dynamics in addition to foliar chemistry and soil mineralogy.



***  


